Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-08-14 measurements.tex
more corrections

%With the testbed, methodology and setups from the previous sections, we
With the methodology and setups from the previous sections, we
proceed to obtain the measurements for the \ac{Raspi} 1 and 2 devices.
We consider the following set of parameters for our study:
For all the codes, we use $g = [16,\ 32,\ 64,\ 128,\ 256,\ 512]$
and $q = [2,\ 2^8]$. For the non-optimized multicore implementations
and cases when the generation size is varied $B = 1600$ bytes.
We consider a setup where we only vary the packet size,
$B = [64,\ 128,\ 256,\ 512,\ 1024,\ 2048,\ 4096,\ 8192,\ 16384,\ 32768
,\ 65536,\ 131072]$ bytes with a generation size fixed on $g = [16,\ 128]$
to see computing load that the \ac{Raspi} present in the low and high
packet size regimes. We considered $B = 1536$ bytes for the optimized
multicore implementation. For \ac{SRLNC}, we considered
$d = [0.02,\ 0.1]$. For \ac{TSNC}, we considered excess packets,
$e = [8,\ 16]$, so that the budget is simply $b = g + e$. In all our
measurement reports, to simplify their review, we first presents the
results for the \ac{Raspi} 1 and later continue with the \ac{Raspi} 2.

@@

\subsubsection{Encoding}
Fig. \ref{fig:enc_good_rasp1} shows the results of the encoder goodput
measurements for the \ac{Raspi} 1. Fig. \ref{fig:enc_good_rasp1_gen_gf2}
and \ref{fig:enc_good_rasp1_gen_gf256} present goodput as function of the
generation size for $GF(2)$ and $GF(2^8)$ respectively for a packet size
of $B = 1600$ bytes. Fig. \ref{fig:enc_good_rasp1_packet_gf2} and
\ref{fig:enc_good_rasp1_packet_gf256} present the same metric but now as
function of the packet size for $GF(2)$ and $GF(2^8)$ respectively. In this
case the generation size was set to $g = 16$ packets.

@@

In the sets of figures shown in Fig.~\ref{fig:enc_good_rasp1} and
\ref{fig:enc_good_rasp2}, it can be observed that the goodput for $GF(2)$
is higher than for $GF(2^8)$ but still around the same order of magnitud.
This difference is explained by noticing that \ac{GF} arithmetics in the
binary field are simply XOR or AND operations. These operations are
implemented efficiently by most of the architectures nowday and are
tailored to be fast in principle. However, the operations in $GF(2^8)$
are more complex given that the finite field arithmetics have to be
performed with lookup tables which at the end reduces the computing
speed giving a lower goodput when compared with the binary field.

In Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen the goodput trends of five codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$ as the number of density changes in the \ac{TSNC} transmission
process, if any. The maximum number of possible changes will depend
on the budget. The larger the budget, the more regions are formed and
more density changes are possible. We first evaluate \ac{TSNC} with no
changes, e.g. $C = 0$ and the two amounts for dependent packets mentioned
earlier. This means that we tolerate at most 8 or 16 packets regardless of
the generation. The goal with setting first \ac{TSNC} with these values
is to observe first the sole effect of the budget in the code since
$C = 0$ implies no density changes during the whole process.

For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case. Given that $C=0$,
this permits it to use a single very low code density $d$. For \ac{RLNC},
this is not the case since the occurrence of \ac{l.d.} coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \cite{trullols2011exact,zhao2012notes},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and \ac{l.d.} coded
packets transmission overhead.

For Fig.~\ref{fig:enc_good_rasp1_packet_gf2},
\ref{fig:enc_good_rasp1_packet_gf256}, \ref{fig:enc_good_rasp2_packet_gf2}
and \ref{fig:enc_good_rasp2_packet_gf256}; we see the packet
size effect in the encoding goodput in both \ac{Raspi} models and for
a fixed generation size. For \ac{TSNC} in this case, we allow for five density
changes during the generation transmission and again consider the same
budgets as before. In all the cases, we observe that there is an
optimal packet size to operate with. When reviewing the specifications
of the \ac{Raspi} 1, it uses a L1 cache of 16 KiB. Hence, the trends
in the packet size can be explained in the following way: For low packet
sizes, the data processing does not overload the \ac{CPU} of the
\ac{Raspi}, so the goodput progressively increases given that we simply
process more data as the packet size increases. However, after a certain
packet size, the goodput gets affected since the cache starts to saturate.
The packets towards the cache needs to have more processing. Beyond this
critical packet size, the \ac{CPU} just queues processing which incurs in
larger delay reducing the goodput.

@@

\subsubsection{Decoding}
Similar as the encoding goodput, in this Section, we review decoder
goodput in terms of performance and configurations.
Fig.~\ref{fig:dec_good_rasp1} shows the results of the decoder goodput
measurements for the \ac{Raspi} 1 and Fig. \ref{fig:dec_good_rasp2} does
the proper for \ac{Raspi} 2.

@@

In Fig.~\ref{fig:dec_good_rasp1} and \ref{fig:dec_good_rasp2}, we
observe the same generation and packet size effects reported in the
encoding case. However, we do observe in
Fig.~\ref{fig:dec_good_rasp1_gen_gf2}, \ref{fig:dec_good_rasp1_gen_gf256},
\ref{fig:dec_good_rasp2_gen_gf2} and \ref{fig:dec_good_rasp2_gen_gf256};
that duplicating the generation size does not reduce by a factor of four the
goodput. In principle, given that Gaussian elimination scales as
$\mathcal{O}(g^3)$ (and thus the processing time), we would expect that
the goodput scales as $\mathcal{O}(R_{proc,dec}) = \mathcal{O}(g/g^{3}) =
\mathcal{O}(g^{-2})$. This would imply that duplicating the generation
size, should reduce the goodput by a factor of 4, which does not. Instead,
the goodput it is only reduced by a factor of 2. This is only possible if
the Gaussian elimination is $\mathcal{O}(g^2)$. A study in
\cite{paramanathan2013leanandmean} for \ac{RLNC} speeds in commercial
devices indicated that this is effectively the case. The reason is that
the $g^2$ scaling factor in the scaling law for our Gaussian elimination is much
higher than the $g^3$ scaling for $g < 512$. Particularly, this factor
relates to the number of field elements in a packet size as mentioned
in \cite{paramanathan2013leanandmean}.

Another difference with the encoding goodput results in the same figures,
is that even though the \ac{TSNC} with $e = 16$ packets provides the fastest
encoding, it does not happen to be the same for the decoding. For this
very sparse scheme, the decoding is affected by the amount of \ac{l.d.}
packets generated by, which leads to a higher delay in some cases
(particularly in $g = [64,\ 128]$). For other generation sizes, the
performance of sparse codes is similar.
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_decoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:dec_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_decoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:dec_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_decoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:dec_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_decoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:dec_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Decoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:dec_good_rasp2}
\end{figure*}
%
\subsection{Average Power}
With the energy measurement setup described in Section
\ref{sec:energy_metrics_methods}, we compute the average power of each
device from their elecric current consumption across time in
Fig.~\ref{fig:current_rasp1} and \ref{fig:current_rasp2}. In each
figure, the eletric current values are classified into three possible
currents: Idle, transition and processing. These values are identified by
our post-processing. Later, the values in red are idle currents, green for
transitions and blue for processing.

@@

\hline
\ac{Raspi} Model & $I_{idle,avg}$ (A)& $I_{proc,avg}$ (A) & $P_{avg}$ (W) \\
\hline
\hline
    \ac{Raspi} 1 &      0.320        &            0.360    & 0.200\\
\hline
    \ac{Raspi} 2 &      0.216        &            0.285    & 0.345\\
\hline
\end{tabular}
\vspace{0.2cm}
\label{tab:average_power_table}
\end{table}
%
From Table~\ref{tab:average_power_table}, we notice that the power expenditure
for both models is almost the same. Thus, the energy behaviour is mostly
dependent on the goodput trends since the power is just a scaling constant.
%
\subsection{Energy per bit}
With power and goodput measurements, we compute the energy per bit of
the previously mentioned cases as described in \eqref{eq:energy_per_bit}.
The trends of the energy per bit are simply the inverse of the goodput since
both metrics are related by an inverse law. As made with the goodput, we
separate the results descriptions according to the operation carried by the
\ac{Raspi}: Encoding and decoding. Besides this, we differentiate between
the models in our study. The energy per bit consumption removes the dependency
on the amount of packets helping to normalize the results and indicating
energy consumption on a fair basis for all the configurations and coding
operations involved in the study.

@@

To review the performance of \ac{NC} in a multicore architecture using the
\ac{Raspi}, we implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the \ac{Raspi} 2 model B, which
features four ARMv7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz clock.
Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The cores
share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with
NEON enabled.
%NEON-enabled code adapted from the Fifi library \cite{fifi}.
The \ac{Raspi} 2 has a NEON extension instruction set which
provides an 128-bit \ac{SIMD} instructions that speed the computations.
Fig.~\ref{enc_dec1024}, \ref{enc_dec128} and \ref{enc_dec16} shows the
encoding and decoding goodput in MB per second for different generation sizes,
$g = [1024, 128, 16]$ respectively. For $g = [128, 16]$, the displayed results
are the mean values over 1000 measurements, while for $g=1024$, they are mean
values over 100 measurements. The size of each coded packet was fixed to
$1536$~bytes since that is the typical size of an Ethernet frame. The blocked
operations were performed dividing the matrices in squared sub-blocks of $16,\
32\ ,\ 64,\ldots,\ 1024$ operands (words in the Galois field) in height and
width. The figures show only block sizes of $16 \times 16$, and $128 \times 128$
operands since with bigger block sizes, the operands do not fit in cache.
Several testcases are considered and detailed.

>>>>

%With the testbed, methodology and setups from the previous sections, we
With the methodology and setups from the previous sections, we
proceed to obtain the measurements for the \ac{Raspi} 1 and 2 devices.
We consider the following set of parameters for our study:
For all the codes, we use $g = [16,\ 32,\ 64,\ 128,\ 256,\ 512]$
and $q = [2,\ 2^8]$. For the non-optimized multicore implementations
and cases when the generation size is varied $B = 1600$ bytes.
We consider a setup where we only vary the packet size,
$B = [64,\ 128,\ 256,\ 512,\ 1024,\ 2048,\ 4096,\ 8192,\ 16384,\ 32768
,\ 65536,\ 131072]$ bytes with a generation size fixed on $g = [16,\ 128]$
to see the performance of the \ac{Raspi}s in low and high
packet size regimes. We considered $B = 1536$ bytes for the optimized
multicore implementation. For \ac{SRLNC}, we considered
$d = [0.02,\ 0.1]$. For \ac{TSNC}, we considered excess packets,
$e = [8,\ 16]$, so that the budget is $b = g + e$. In all our
measurement reports, to simplify their review, we first presents the
results for the \ac{Raspi} 1 and later continue with the \ac{Raspi} 2.

@@

\subsubsection{Encoding}
Fig. \ref{fig:enc_good_rasp1} shows the results of the encoder goodput
measurements for the \ac{Raspi} 1. Fig. \ref{fig:enc_good_rasp1_gen_gf2}
and \ref{fig:enc_good_rasp1_gen_gf256} presents goodput as function of the
generation size for $GF(2)$ and $GF(2^8)$ respectively for a packet size
of $B = 1600$ bytes. Fig. \ref{fig:enc_good_rasp1_packet_gf2} and
\ref{fig:enc_good_rasp1_packet_gf256} present the same metric but now as
function of the packet size for $GF(2)$ and $GF(2^8)$ respectively. In this
case the generation size was set to $g = 16$ packets.

@@

In the sets of figures shown in Fig.~\ref{fig:enc_good_rasp1} and
\ref{fig:enc_good_rasp2}, it can be observed that the goodput for $GF(2)$
is higher than for $GF(2^8)$ but still around the same order of magnitude.
This difference is explained by noticing that \ac{GF} arithmetics in the
binary field are simply XOR or AND operations. These operations are
implemented efficiently by most of the architectures nowadays and are
tailored to be fast in principle. However, the operations in $GF(2^8)$
are more complex given that the finite field arithmetics have to be
performed with lookup tables which at the end reduces the computing
speed giving a lower goodput when compared with the binary field.

In Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, \ref{fig:enc_good_rasp1_gen_gf256},
\ref{fig:enc_good_rasp2_gen_gf2} and \ref{fig:enc_good_rasp2_gen_gf256};
it can be seen that the goodput trends of five codes as a function of
the generation size: \ac{RLNC}, \ac{SRLNC} with $d = [0.02,\ 0.1]$
and \ac{TSNC} with an extra parameter that we mention. We define
$C$\fxnote{C is already defined as the set of l.i. coded packets}
as the number of density regions in the \ac{TSNC} transmission
process. The maximum number of possible regions depends
on the generation size. See equation~\eqref{eq:feedback_dofs}.
The larger the generation, the more density regions may
be formed and more density changes are possible.
Throughout this paper, \ac{TSNC} is configured to use the maximum number
of density regions possible. 
As there can be at least $C=1$ density region in a transmission, we use
$C=0$ to indicate the maximum possible density regions. This is
used for \ac{TSNC} in plots where the generation size is not fixed.
%In fact, \ac{TSNC} is configured to use the
%maximum possible density regions in all plots throughout this paper.
%The larger the generation, the more regions are formed and
%more density changes are possible.
%Throughout the section, we evaluate \ac{TSNC} with as many possible
%density regions that are possible to form during a transmission. This number
%grows as the generation increases, so we define $C=0$ as the maximum
%number of possible regions.
%
%We first evaluate \ac{TSNC} with no
%changes, e.g. $C = 0$ and the two amounts for dependent packets mentioned
%earlier. This means that we tolerate at most 8 or 16\fxnote{dependent packets} packets regardless of
%the generation. The goal with setting first \ac{TSNC} with these values
%is to observe first the sole effect of the budget in the code since
%$C = 0$ implies no density changes during the whole process.
%
For the plots in the mentioned figures from the encoder goodput measurements,
\ac{RLNC} presents the lowest performance in terms of goodput and
\ac{TSNC} with $e = 16$, the highest regardless of the \ac{Raspi} model.
Given that the processing time depends in the amount of coded packets
required to create, \ac{RLNC} is the slowest to process since it must
use all of the $g$ original packets\fxnote{The encoding process speed depends on the total number of original packets to mix. Thus coding density and number of coded packets per generation}. Later, sparse codes process the data
at a larger rate since less packets are being mixed when creating a
coded packet. The caveat of these schemes is that the sparser the code,
the more probable the ocurrences of \ac{l.d.} packets is. So, basically
the sparser the codes, then more overhead due to transmissions of \ac{l.d.}
packets since it is more frequent to generate similar set of packets.
Excluding the coding coefficients overhead, the overhead due to
transmissions of \ac{l.d.} packets might be high for the sparse schemes.
For example, if we consider \ac{TSNC} with $e = 16$ and $g = 16$
in Fig.~\ref{fig:enc_good_rasp1_gen_gf2}, the budget in this case permits
to send up to 32 packets which is $2\times$ the generation
size for an overhead of $100\%$ excluding the overhead from appending
the coding coefficients. This happens because \ac{TSNC}
has been allowed to add too much redundancy in this case. Given that $C=0$,
this permits it to use a single very low code density $d$. For \ac{RLNC},
this is not the case since the occurrence of \ac{l.d.} coded packets is
small, because all coding coefficients are used. Even for $GF(2)$, the
average amount of redundant packets for \ac{RLNC} has proven to be 1.6
packets after $g$ have been transmitted \cite{trullols2011exact,zhao2012notes},
but less than the cases where sparse codes are utilized. Overall, we
observe that there is a trade-off between goodput and \ac{l.d.} coded
packets transmission overhead.

For Fig.~\ref{fig:enc_good_rasp1_packet_gf2},
\ref{fig:enc_good_rasp1_packet_gf256}, \ref{fig:enc_good_rasp2_packet_gf2}
and \ref{fig:enc_good_rasp2_packet_gf256}; we see the packet
size effect in the encoding goodput in both \ac{Raspi} models and for
a fixed generation size. For \ac{TSNC} in this case, we allow for five density
changes during the generation transmission and again consider the same
budgets as before. In all the cases, we observe that there is an
optimal packet size to operate with. When reviewing the specifications
of the \ac{Raspi} 1, it uses a L1 cache of 16 KiB. Hence, the trends
in the packet size can be explained in the following way: For low packet
sizes, the data processing does not overload the \ac{CPU} of the
\ac{Raspi}, so the goodput progressively increases given that we
process more data as the packet size increases. However, after a certain
packet size, the goodput gets affected since the cache starts to saturate.
The packets towards the cache needs to have more processing. Beyond this
critical packet size, the \ac{CPU} just queues processing which incurs in
larger delay reducing the goodput.

@@

\subsubsection{Decoding}
Similar as the encoding goodput, in this Section, we review decoder
goodput in terms of performance and configurations.
Fig.~\ref{fig:dec_good_rasp1} shows the results of the decoder goodput
measurements for the \ac{Raspi} 1 and Fig. \ref{fig:dec_good_rasp2} for the \ac{Raspi} 2.

@@

In Fig.~\ref{fig:dec_good_rasp1} and \ref{fig:dec_good_rasp2}, we
observe the same generation and packet size effects reported in the
encoding case. However, we do observe in
Fig.~\ref{fig:dec_good_rasp1_gen_gf2}, \ref{fig:dec_good_rasp1_gen_gf256},
\ref{fig:dec_good_rasp2_gen_gf2} and \ref{fig:dec_good_rasp2_gen_gf256};
that duplicating\fxnote{What does duplicating the generation size mean? Did you mean doubling?} the generation size does not reduce the goodput by a
factor of four. In principle, given that Gaussian elimination scales as
$\mathcal{O}(g^3)$ (and thus the processing time), we would expect that
the goodput to scale as $\mathcal{O}(R_{proc,dec}) = \mathcal{O}(g/g^{3}) =
\mathcal{O}(g^{-2})$. This would imply that duplicating\fxnote{doubling?} the generation
size, should reduce the goodput by a factor of four, which is not the case. Instead,
the goodput is only reduced by a factor of two. This is only possible if
the Gaussian elimination is $\mathcal{O}(g^2)$. A study in
\cite{paramanathan2013leanandmean} for \ac{RLNC} speeds in commercial
devices indicated that this is effectively the case. The reason is that
the $g^2$ scaling factor in the scaling law\fxnote{scaling estimate or scaling expression or something else instead of scaling law} of the Gaussian elimination
is much
higher than the $g^3$ scaling for $g < 512$. Particularly, this factor
relates to the number of field elements in a packet size as mentioned
in \cite{paramanathan2013leanandmean}.

Another difference with the encoding goodput results in the same figures,
is that even though \ac{TSNC} with $e = 16$ packets provides the fastest
encoding, it does not happen to be the same for the decoding. For this
very sparse scheme, the decoding is affected by the amount of \ac{l.d.}
packets generated by, which leads to a higher delay in some cases
(particularly in $g = [64,\ 128]$). For other generation sizes, the
performance of sparse codes is similar.\fxnote{This block is incorrect. e=16 is not the fastest encoder nor decoder. e=8 is!}
%
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_decoder_Binary_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2$}}
        \label{fig:dec_good_rasp2_gen_gf2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_generation_size_Rasp_v2_decoder_Binary8_1600.eps}
        \caption[]%
        {{\small Goodput vs. Generation size for $q = 2^8$}}
        \label{fig:dec_good_rasp2_gen_gf256}
    \end{subfigure}
    \vskip\baselineskip
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_decoder_Binary_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2,\ g=16$}}
        \label{fig:dec_good_rasp2_packet_gf2}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=1.15\textwidth]{images/06_06_2016/goodput_vs_symbol_size_Rasp_v2_decoder_Binary8_16.eps}
        \caption[]%
        {{\small Goodput vs. Packet size for $q = 2^8,\ g=16$}}
        \label{fig:dec_good_rasp2_packet_gf256}
    \end{subfigure}
    \caption[]
    {\small Decoder goodput measurements for the \ac{Raspi} 2}
    \label{fig:dec_good_rasp2}
\end{figure*}
%
\subsection{Average Power}
With the energy measurement setup described in Section
\ref{sec:energy_metrics_methods}, we compute the average power of each
device from their elecric current consumption across time in
Fig.~\ref{fig:current_rasp1} and \ref{fig:current_rasp2}. In each
figure, the eletric current values are classified into three possible
currents: Idle, transition and processing. These values are identified by
our post-processing. Later, the values in red are idle currents, green for
transitions and blue for processing.

@@

\hline
\ac{Raspi} Model & $I_{idle,avg}$ (A)& $I_{proc,avg}$ (A) & $P_{avg}$ (W) \\
\hline
\hline
    \ac{Raspi} 1 &      0.320        &            0.360    & 0.200\\
\hline
    \ac{Raspi} 2 &      0.216        &            0.285    & 0.345\\
\hline
\end{tabular}
\vspace{0.2cm}
\label{tab:average_power_table}
\end{table}
%
From Table~\ref{tab:average_power_table}, we notice that the power expenditure
for both models is almost the same. Thus, the energy behaviour is mostly
dependent on the goodput trends since the power is just a scaling constant.
%
\subsection{Energy per bit}
With power and goodput measurements, we compute the energy per bit of
the previously mentioned cases as described in \eqref{eq:energy_per_bit}.
The trends of the energy per bit are simply\fxnote{The word "simply" should be forbidden} the inverse of the goodput since
both metrics are related by an inverse law. As made with the goodput, we
separate the results descriptions according to the operation carried by the
\ac{Raspi}: Encoding and decoding. Besides this, we differentiate between
the models in our study. The energy per bit consumption removes the dependency
on the amount of packets helping to normalize the results and indicating
energy consumption on a fair basis for all the configurations and coding
operations involved in the study.

@@

To review the performance of \ac{NC} in a multicore architecture,
%using the \ac{Raspi}
we implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the \ac{Raspi} 2 model B, which
features four ARMv7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz clock.\fxnote{... which features an ARMv7 CPU with four cores in a Broadcom ...}
Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The cores
share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with
NEON enabled.
%NEON-enabled code adapted from the Fifi library \cite{fifi}.
The \ac{Raspi} 2 has a NEON extension instruction set which
provides an 128-bit \ac{SIMD} instructions that speed the computations.
Fig.~\ref{enc_dec1024}, \ref{enc_dec128} and \ref{enc_dec16} shows the
encoding and decoding goodput in MB per second for different generation sizes,
$g = [1024, 128, 16]$ respectively. For $g = [128, 16]$, the displayed results
are the mean values over 1000 measurements, while for $g=1024$, they are mean
values over 100 measurements. The size of each coded packet was fixed to
$1536$~bytes since that is the typical size of an Ethernet frame. The blocked
operations were performed dividing the matrices in squared sub-blocks of $16,\
32\ ,\ 64,\ldots,\ 1024$ operands (words in the Galois field) in height and
width. The figures show only block sizes of $16 \times 16$, and $128 \times 128$
operands since with bigger block sizes, the operands do not fit in cache.
Several testcases are considered and detailed.

### 2016-08-14 metrics.tex
more corrections

\subsubsection{Average Power Expenditure}
To extract the average power for a given configuration in our setup, we
first calculate the average current in each the idle state $I_{idle,avg}$
and the processing state $I_{proc,avg}$ for all the available sample for
the given configuration. Regardless of the current type, the average value
of the sample set of a given number of samples $N_s$ is:
%
\begin{align} \label{eq:avg_current}
I_{avg} = \frac{1}{N_s}\sum_{k=1}^{N_s} I_{k}  ~[\mathrm{Ampere}]
\end{align}
%
With the average current from \eqref{eq:avg_current}, we compute the average
current used for processing with respect to the idle state, by substracting
$I_{idle,avg}$ from $I_{proc,avg}$. Then, the result is multiplied by the
supply voltage to obtain the average power during the considered configuration,
given as:
%
\begin{align} \label{eq:avg_power}
P_{avg} = V_{supply}(I_{proc,avg} - I_{idle,avg}) ~[\mathrm{Watt}]
\end{align}
%
\subsubsection{Energy per Bit Consumption}
To get this metric for a given configuration, we simply express the energy
as the product of the average power by the processing time $T_{proc}
~[\mathrm{sec/byte}]$ obtained from the goodput measurement for the same
configuration. In this way, we can relate the processing time with the
goodput, the packet and the generation size as shown:
%
\begin{align} \label{eq:energy_per_bit}
E_b = P_{avg} T_{proc,bit} = P_{avg} \times \frac{T_{proc}}{8gB} = \frac{P_{avg}}{8R_{proc}} ~[\mathrm{Joule}]
\end{align}
%

>>>>

\subsubsection{Average Power Expenditure}
To extract the average power for a given configuration in our setup, we
first calculate the average current in each\fxnote{each of?} the idle state $I_{idle,avg}$
and the processing state $I_{proc,avg}$ for all the available sample for
the given configuration. Regardless of the current type, the average value
of the sample set of a given number of samples $N_s$ is:
%
\begin{align} \label{eq:avg_current}
I_{avg} = \frac{1}{N_s}\sum_{k=1}^{N_s} I_{k}  ~[\mathrm{Ampere}]
\end{align}
%
With the average current from \eqref{eq:avg_current}, we compute the average
current used for processing with respect to the idle state, by substracting
$I_{idle,avg}$ from $I_{proc,avg}$. Then, the result is multiplied by the
supply voltage to obtain the average power during the considered configuration,
given as:
%
\begin{align} \label{eq:avg_power}
P_{avg} = V_{supply}(I_{proc,avg} - I_{idle,avg}) ~[\mathrm{Watt}]
\end{align}
%
\subsubsection{Energy per Bit Consumption}
To get this metric for a given configuration, we express the energy
as the product of the average power by the processing time $T_{proc}
~[\mathrm{sec/byte}]$ obtained from the goodput measurement for the same
configuration. In this way, we can relate the processing time with the
goodput, the packet and the generation size as shown:
%
\begin{align} \label{eq:energy_per_bit}
E_b = P_{avg} T_{proc,bit} = P_{avg} \times \frac{T_{proc}}{8gB} = \frac{P_{avg}}{8R_{proc}} ~[\mathrm{Joule}]
\end{align}
%
